Towards human SuperEEG

### 2018-10-11 paper/_region_.tex
edits to paper and figures

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from 0.3--14.2 hours
(Fig.~\suppstats A), and includes data recorded roughly from when the
patients woke up each morning, to before they went to sleep at the end
of each day.  In addition to typical bed-ridden hospital patient
activities (e.g., lying in bed, reading, watching television, using
personal electronic devices, listening to music, visiting with family
and friends, etc.), the patients also performed a variety of
experimental cognitive tasks throughout their day (primarily
list-learning memory tasks).  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from 3159 electrodes implanted in the brains
of 24 epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from .4--6.6 hours
(Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

@@

To illustrate our approach, we first examine an individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (red), and its associated
reconstruction (blue), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = 0.869, p = 10\textsuperscript{-10]$).  Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example
electrode (correlation: $r = 0.909, p = 10\textsuperscript{-10}$).  This example confirms
that the SuperEEG algorithm recovers the recordings from this single
electrode well.  Next, we used this general approach to quantify the
algorithm's performance across the full dataset.

%sample trace & reconstruction, several frames of activity snapshot?
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/recon}
  \caption{\textbf{Observed and reconstructed LFP from a single
      electrode.} \textbf{A. Example LFP.}  A 5~s recording from the
    red electrode in Figure~\ref{fig:methods}A is displayed in red,
    and the reconstructed LFP during the same time window is shown in
    blue.  \textbf{B. Observed versus reconstructed LFP over
      14.2~hours.}  The two-dimensional histogram reflects the
    relation between distributions of observed versus reconstructed
    voltages from one patient, across the 14.2 hours of recorded data
    collected over 6 recording sessions.  The correlation reported in
    the panel is between the observed and reconstructed voltages.
    Both panels: all voltages are represented in standard deviation units
   (computed within session).}
  \label{fig:recon}
\end{figure}

For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: XXX; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(XXX) = XXX, p = XXX$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model learned from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from that single patient may be used to estimate held-out data from
the same patient.  This allows us to ask how much information about
the activity at a given electrode might be inferred through (a) volume
conductance or other sources of ``leakage'' from activity patterns
measured from the patient's other electrodes and (b) across-electrode
correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: XXX;
$t$-test comparing mean of distribution of $z$-transformed
per-electrode correlation coefficients to 0: $t(XXX) = XXX, p = XXX$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(XXX) = XXX, p = XXX$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent dataset and found
similar results (Fig.~\ref{fig:corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). We also
replicated this result separately for each of the two experiments from
Dataset 2 (Fig.~\perexpcorrmaps).  This overall finding, that
reconstructions of held-out data using correlation models learned from
\textit{other} patient's data yield higher reconstruction accuracy
than correlation models learned from the patient whose data is being
reconstructed, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beyond the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

To test the task-specificity of the SuperEEG-derived correlation
models, we repeated the above within- and across-patient cross validation
procedures separately for Experiment 1 and Experiment 2 data from
Dataset 2.  We then compared the reconstruction accuracies for held-out
electrodes, for models trained within versus across the two
experiments, or combining across both experiments
(Fig.~\perexptaskrecon).  In every case we found that across-patient
models trained using data from all other patients out-performed
within-patient models trained on data only from the subject
contributing the given electrode ($t$s$(XXX) > XXX, p$s$ < XXX$).  All
reconstruction accuracies also reliably exceeded chance performance
($t$s$(XXX) > XXX, p$s$ < XXX$).  Average reconstruction accuracy was
highest for the across-patient models limited to data from the same
experiment (mean accuracy: $0.55$); next-highest for the
across-patient models that combined data across both experiments (mean
accuracy: $0.50$); and lowest for models trained across task (mean
accuracy: $0.37$).  This result also held for each of the Dataset 2
experiments individually (Fig.~\perexptaskreconseparated).  Taken
together, these results indicate that there are reliable commonalities
in the spatial correlations of full-brain activity across tasks, but
that there are also reliable differences in these spatial correlations
across tasks.  Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is
highest when constrained to within-task data, or data that includes a
variety of tasks (e.g., Dataset 1, or combining across the two Dataset
2 experiments).

Although both datasets we examined provide good full-brain coverage
(when considering data from every patient; e.g.\
Fig.~\ref{fig:corrmap}C, D), electrodes are not placed uniformly
throughout the brain.  For example, electrodes are more likely to be
implanted in regions like the medial temporal lobe (MTL), and are
rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel
MNI152 brain, we computed the proportion of electrodes in the dataset
that were contained within a 20 MNI unit radius sphere centered on
that voxel.  We defined the \textit{density} at that location as this
proportion.  Across Datasets 1 and 2, the electrode placement
densities were similar (correlation by voxel:
$r = XXX, p = XXX$).  We wondered whether regions with good
covererage might be associated with better reconstruction accuracy
(e.g. Fig.~\ref{fig:corrmap}C, D indicate that many electrodes in the
MTL have relatively high reconstruction accuracy, and occipital
electrodes tend to have relatively low reconstruction accuracy).  To
test whether this held more generally across the entire brain, for
each dataset we computed the electrode placement density for each
electrode from each patient (using the proportion of \textit{other}
patients' electrodes within 20 MNI units of the given electrode).  We
then correlated these density values with the across-patient
reconstruction accuracies for each electrode.  Contrary to our
expectation, rather than positive correlations, we found weak (but
reliable) negative correlations between reconstruction accuracy and
density for both datasets (Dataset 1: $r = XXX, p = XXX$; Dataset 2:
$r = -0.16, p < 10^{-10}$).  This indicates that the reconstruction
accuracies we observed are not driven solely by sampling density, but
rather may also reflect higher order properties of neural dynamics
such as functional correlations between distant
voxels~\citep{BetzEtal17b}.

@@

\message{ !name(main.tex) !offset(-850) }

>>>>

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from {0.3}--14.2 hours
(Fig.~\suppstats A), and includes data from a list-learning memory experiment
for up to one hour varying by patient with additional buffer time in
which the patients performed typical bed-ridden hospital
activities (e.g., lying in bed, reading, watching television, using
personal electronic devices, listening to music, visiting with family
and friends, etc.).  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from 4436 electrodes implanted in the brains
of 40 epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from {0.4}--6.6 hours
(Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

@@

To illustrate our approach, we first examine an individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (blue), and its associated
reconstruction (red), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = 0.869, p < 10^{-10}$).  Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example
electrode (correlation: $r = 0.909, p < 10^{-10}$).  This example confirms
that the SuperEEG algorithm recovers the recordings from this single
electrode well.  Next, we used this general approach to quantify the
algorithm's performance across the full dataset.

%sample trace & reconstruction, several frames of activity snapshot?
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/recon}
  \caption{\textbf{Observed and reconstructed LFP from a single
      electrode.} \textbf{A. Example LFP.}  A 5~s recording from the
    red electrode in Figure~\ref{fig:methods}A is displayed in blue,
    and the reconstructed LFP during the same time window is shown in
    red.  \textbf{B. Observed versus reconstructed LFP over
      14.2~hours.}  The two-dimensional histogram reflects the
    relation between distributions of observed versus reconstructed
    voltages from one patient, across the 14.2 hours of recorded data
    collected over 6 recording sessions.  The correlation reported in
    the panel is between the observed and reconstructed voltages.
    Both panels: all voltages are represented in standard deviation units
   (computed within session).}
  \label{fig:recon}
\end{figure}

For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: 0.52; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(66) = 25.08, p < 10^{-30}$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model learned from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from that single patient may be used to estimate held-out data from
the same patient.  This allows us to ask how much information about
the activity at a given electrode might be inferred through (a) volume
conductance or other sources of ``leakage'' from activity patterns
measured from the patient's other electrodes and (b) across-electrode
correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: 0.32;
$t$-test comparing mean of distribution of $z$-transformed
per-electrode correlation coefficients to 0: $t(66) = 15.16, p < 10^{-20}$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(66) = 9.62, p < 10^{-10}$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent dataset and found
similar results (Fig.~\ref{fig:corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). We also
replicated this result separately for each of the two experiments from
Dataset 2 (Fig.~\perexpcorrmaps).  This overall finding, that
reconstructions of held-out data using correlation models learned from
\textit{other} patient's data yield higher reconstruction accuracy
than correlation models learned from the patient whose data is being
reconstructed, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beyond the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

To test the task-specificity of the SuperEEG-derived correlation
models, we repeated the above within- and across-patient cross validation
procedures separately for Experiment 1 and Experiment 2 data from
Dataset 2.  We then compared the reconstruction accuracies for held-out
electrodes, for models trained within versus across the two
experiments, or combining across both experiments
(Fig.~\perexptaskrecon).  In every case we found that across-patient
models trained using data from all other patients out-performed
within-patient models trained on data only from the subject
contributing the given electrode ($t$s$(24) > 6.50, p$s$ < 10^{-5}$).  All
reconstruction accuracies also reliably exceeded chance performance
($t$s$(24) > 8.00, p$s$ < 10^{-8}$).  Average reconstruction accuracy was
highest for the across-patient models limited to data from the same
experiment (mean accuracy: $0.68$); next-highest for the
across-patient models that combined data across both experiments (mean
accuracy: $0.61$); and lowest for models trained across task (mean
accuracy: $0.47$).  This result also held for each of the Dataset 2
experiments individually (Fig.~\perexptaskreconseparated).  Taken
together, these results indicate that there are reliable commonalities
in the spatial correlations of full-brain activity across tasks, but
that there are also reliable differences in these spatial correlations
across tasks.  Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is
highest when constrained to within-task data, or data that includes a
variety of tasks (e.g., Dataset 1, or combining across the two Dataset
2 experiments).

Although both datasets we examined provide good full-brain coverage
(when considering data from every patient; e.g.\
Fig.~\ref{fig:corrmap}C, D), electrodes are not placed uniformly
throughout the brain.  For example, electrodes are more likely to be
implanted in regions like the medial temporal lobe (MTL), and are
rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel
MNI152 brain, we computed the proportion of electrodes in the dataset
that were contained within a 20 MNI unit radius sphere centered on
that voxel.  We defined the \textit{density} at that location as this
proportion.  Across Datasets 1 and 2, the electrode placement
densities were similar (correlation by voxel:
$r = 0.56, p < 10^{-10}$).  We wondered whether regions with good
covererage might be associated with better reconstruction accuracy
(e.g. Fig.~\ref{fig:corrmap}C, D indicate that many electrodes in the
MTL have relatively high reconstruction accuracy, and occipital
electrodes tend to have relatively low reconstruction accuracy).  To
test whether this held more generally across the entire brain, for
each dataset we computed the electrode placement density for each
electrode from each patient (using the proportion of \textit{other}
patients' electrodes within 20 MNI units of the given electrode).  We
then correlated these density values with the across-patient
reconstruction accuracies for each electrode.  Contrary to our
expectation, rather than positive correlations, we found weak (but
reliable) negative correlations between reconstruction accuracy and
density for both datasets (Dataset 1: $r = -0.07, p < 10^{-5}$; Dataset 2:
$r = -0.18, p < 10^{-10}$).  This indicates that the reconstruction
accuracies we observed are not driven solely by sampling density, but
rather may also reflect higher order properties of neural dynamics
such as functional correlations between distant
voxels~\citep{BetzEtal17b}.

@@

\message{ !name(main.tex) !offset(-849) }
